Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Shanghai version \YellowPaperVersionNumber}}

### 2025-02-26 Paper.tex
typo

Note we use $\hyperlink{theta}{\Theta}_{4}$ and $\hyperlink{lambda}{\Lambda}_{4}$ to denote the fact that only the first four components of the functions' values are taken; the finals represents the message-call's output value (a byte array), and the new transient storage and are unused in the context of transaction evaluation.

@@

\begin{equation}
(\boldsymbol{\sigma}^{**}, \boldsymbol{\theta}^{**}, g^{**}, A^{**}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, \boldsymbol{\theta}^*, g, A^*, I) \\
\end{equation}
\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
\begin{eqnarray}
I_{\mathrm{a}} & \equiv & a \\
I_{\mathrm{o}} & \equiv & o \\
I_{\mathrm{p}} & \equiv & p \\
I_{\mathbf{d}} & \equiv & () \\
I_{\mathrm{s}} & \equiv & s \\
\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
I_{\mathbf{b}} & \equiv & \mathbf{i} \\
I_{\mathrm{e}} & \equiv & e \\
I_{\mathrm{w}} & \equiv & w
\end{eqnarray}

@@

\begin{eqnarray}
\boldsymbol{\sigma}' & \equiv & \begin{cases}
\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
\boldsymbol{\sigma}^{**} & \text{otherwise}
\end{cases} \\
\boldsymbol{\theta}' & \equiv & \begin{cases}
\boldsymbol{\theta} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
\boldsymbol{\theta}^{**} & \text{otherwise}
\end{cases} \\
g' & \equiv & \begin{cases}
0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \\
&\quad \mathbf{o} = \varnothing \\
g^{**} & \text{otherwise}
\end{cases} \\
A' & \equiv & \begin{cases}
A & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
A^{**} & \text{otherwise}
\end{cases} \\
z & \equiv & \begin{cases}
0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
1 & \text{otherwise}
\end{cases}
\end{eqnarray}
\begin{equation}
\hypertarget{code_execution_result}{}
(\boldsymbol{\sigma}^{**}, \boldsymbol{\theta}^{**}, g^{**}, A^{**}, \mathbf{o}) \equiv \Xi
\end{equation}
\begin{eqnarray}
I_{\mathrm{a}} & \equiv & r \\
I_{\mathrm{o}} & \equiv & o \\
I_{\mathrm{p}} & \equiv & p \\
I_{\mathbf{d}} & \equiv & \mathbf{d} \\
I_{\mathrm{s}} & \equiv & s \\
I_{\mathrm{v}} & \equiv & \tilde{v} \\
I_{\mathrm{e}} & \equiv & e \\
I_{\mathrm{w}} & \equiv & w
\end{eqnarray}
\nopagebreak[1]where
\begin{equation}
\Xi \equiv \begin{cases}
\Xi_{\mathtt{ECREC}}    (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 1 \\
\Xi_{\mathtt{SHA256}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 2 \\
\Xi_{\mathtt{RIP160}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 3 \\
\Xi_{\mathtt{ID}}       (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 4 \\
\Xi_{\mathtt{EXPMOD}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 5 \\
\Xi_{\mathtt{BN\_ADD}}  (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 6 \\
\Xi_{\mathtt{BN\_MUL}}  (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 7 \\
\Xi_{\mathtt{SNARKV}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 8 \\
\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 9 \\
\Xi                     (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{otherwise} \end{cases}
\end{equation}
and
\begin{equation}
\mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
\end{equation}
It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.

@@

The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the resultant transient storage $\boldsymbol{\theta}'$, the remaining gas $g'$, the resultant accrued substate $A'$ and the resultant output, $\mathbf{o}$, given these definitions.
For the present context, we will define it as:
\begin{equation}
(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I)
\end{equation}
where we will remember that $A$, the accrued substate, is defined in section \ref{ch:substate}.

@@

\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
\begin{eqnarray}
\Xi(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\theta}', \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}) \\
(\boldsymbol{\sigma}', \boldsymbol{\theta}', \boldsymbol{\mu}'\!, A', ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) \\
\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
\end{eqnarray}
\begin{equation} \label{eq:X-def}
X\big( (\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
\big(\varnothing, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \\
\big(\varnothing, \boldsymbol{\theta}, \boldsymbol{\mu}', A, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
O(\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
X\big(O(\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
\end{cases}
\end{equation}

@@

Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
\begin{eqnarray}
O\big((\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\theta}', \boldsymbol{\mu}', A', I) \\
\Delta & \equiv & \mathbf{\alpha}_{w} - \mathbf{\delta}_{w} \\
\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
\quad \forall x \in [\mathbf{\alpha}_{w}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
\end{eqnarray}

@@

For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
\begin{equation} \label{eq:pre}
\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I) \equiv \begin{cases}
(\varnothing, \boldsymbol{\theta}, 0, A, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
(\boldsymbol\sigma, \boldsymbol{\theta}, g - g_{\mathrm{r}}, A, \mathbf{o}) & \text{otherwise}\end{cases}
\end{equation}

>>>>

Note we use $\hyperlink{theta}{\Theta}_{4}$ and $\hyperlink{lambda}{\Lambda}_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array), and the new transient storage and are unused in the context of transaction evaluation.

@@

\begin{equation}
(\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}, \boldsymbol{\theta}^{**}) \equiv \Xi(\boldsymbol{\sigma}^*, \boldsymbol{\theta}^*, g, A^*, I) \\
\end{equation}

where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:
\begin{eqnarray}
I_{\mathrm{a}} & \equiv & a \\
I_{\mathrm{o}} & \equiv & o \\
I_{\mathrm{p}} & \equiv & p \\
I_{\mathbf{d}} & \equiv & () \\
I_{\mathrm{s}} & \equiv & s \\
\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
I_{\mathbf{b}} & \equiv & \mathbf{i} \\
I_{\mathrm{e}} & \equiv & e \\
I_{\mathrm{w}} & \equiv & w
\end{eqnarray}

@@

\begin{eqnarray}
\boldsymbol{\sigma}' & \equiv & \begin{cases}
\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
\boldsymbol{\sigma}^{**} & \text{otherwise}
\end{cases} \\
\boldsymbol{\theta}' & \equiv & \begin{cases}
\boldsymbol{\theta} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
\boldsymbol{\theta}^{**} & \text{otherwise}
\end{cases} \\
g' & \equiv & \begin{cases}
0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \\
&\quad \mathbf{o} = \varnothing \\
g^{**} & \text{otherwise}
\end{cases} \\
A' & \equiv & \begin{cases}
A & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
A^{**} & \text{otherwise}
\end{cases} \\
z & \equiv & \begin{cases}
0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
1 & \text{otherwise}
\end{cases}
\end{eqnarray}
\begin{equation}
\hypertarget{code_execution_result}{}
(\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}, \boldsymbol{\theta}^{**}) \equiv \Xi
\end{equation}
\begin{eqnarray}
I_{\mathrm{a}} & \equiv & r \\
I_{\mathrm{o}} & \equiv & o \\
I_{\mathrm{p}} & \equiv & p \\
I_{\mathbf{d}} & \equiv & \mathbf{d} \\
I_{\mathrm{s}} & \equiv & s \\
I_{\mathrm{v}} & \equiv & \tilde{v} \\
I_{\mathrm{e}} & \equiv & e \\
I_{\mathrm{w}} & \equiv & w
\end{eqnarray}
\nopagebreak[1]where
\begin{equation}
\Xi \equiv \begin{cases}
\Xi_{\mathtt{ECREC}}    (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 1 \\
\Xi_{\mathtt{SHA256}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 2 \\
\Xi_{\mathtt{RIP160}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 3 \\
\Xi_{\mathtt{ID}}       (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 4 \\
\Xi_{\mathtt{EXPMOD}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 5 \\
\Xi_{\mathtt{BN\_ADD}}  (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 6 \\
\Xi_{\mathtt{BN\_MUL}}  (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 7 \\
\Xi_{\mathtt{SNARKV}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 8 \\
\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 9 \\
\Xi                     (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{otherwise} \end{cases}
\end{equation}
and
\begin{equation}
\mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
\end{equation}
It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.

@@

The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the resultant transient storage $\boldsymbol{\theta}'$, the remaining gas $g'$, the resultant accrued substate $A'$ and the resultant output, $\mathbf{o}$, given these definitions.
For the present context, we will define it as:
\begin{equation}
(\boldsymbol{\sigma}', g', A', \mathbf{o}, \boldsymbol{\theta}') \equiv \Xi(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I)
\end{equation}
where we will remember that $A$, the accrued substate, is defined in section \ref{ch:substate}.

@@

\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
\begin{eqnarray}
\Xi(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}, \boldsymbol{\theta}') \\
(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A', ..., \mathbf{o}, \boldsymbol{\theta}') & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) \\
\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
\end{eqnarray}
\begin{equation} \label{eq:X-def}
X\big( (\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
\big(\varnothing, \boldsymbol{\mu}, A, I, \varnothing, \boldsymbol{\theta}\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \\
\big(\varnothing, \boldsymbol{\mu}', A, I, \mathbf{o}, \boldsymbol{\theta}\big) & \text{if} \quad w = \text{\small REVERT} \\
O(\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
X\big(O(\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
\end{cases}
\end{equation}

@@

Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
\begin{eqnarray}
O\big((\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I, \boldsymbol{\theta}') \\
\Delta & \equiv & \mathbf{\alpha}_{w} - \mathbf{\delta}_{w} \\
\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
\quad \forall x \in [\mathbf{\alpha}_{w}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
\end{eqnarray}

@@

For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
\begin{equation} \label{eq:pre}
\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I) \equiv \begin{cases}
(\varnothing, 0, A, (), \boldsymbol{\theta}) & \text{if} \quad g < g_{\mathrm{r}} \\
(\boldsymbol\sigma, g - g_{\mathrm{r}}, A, \mathbf{o}, \boldsymbol{\theta}) & \text{otherwise}\end{cases}
\end{equation}
